Search Strategies for Functional Logic Programming

### 2012-02-14 papers/atps2012/paper.tex
Minor improvements and additions to ATPS'12 paper

%\IfFileExists{latin1.sty}{\usepackage{latin1}}{\usepackage{isolatin1}}

@@

\author{
Michael Hanus
\quad
Bj{\"o}rn Peem{\"o}ller
\quad
Fabian Reck \\
\\
Institut f\"ur Informatik, CAU Kiel, D-24098 Kiel, Germany \\
\texttt{\{mh|bjp|fre\}@informatik.uni-kiel.de}
}
\title{Search Strategies for Functional Logic Programming}
\begin{document}
\pagestyle{plain} % remove it in the camera-ready version
\maketitle

@@

Functional logic languages combine the most important
features of functional and logic programming in a single language
(see \cite{AntoyHanus10CACM,Hanus07ICLP} for recent surveys).
In particular, they provide higher-order functions and demand-driven
evaluation from functional programming together with logic programming features
like non-deterministic search and computing with partial information
(logic variables).
This combination
led to new design patterns \cite{AntoyHanus02FLOPS,AntoyHanus11WFLP}
and better abstractions for application programming,
e.g., as shown for programming with databases
\cite{BrasselHanusMueller08PADL,Fischer05},
GUI programming \cite{Hanus00PADL},
web programming \cite{Hanus01PADL,Hanus06PPDP,HanusKoschnicke10PADL},
or string parsing \cite{CaballeroLopez99}.
Moreover, it is also a good basis
to teach the ideas of functional and logic programming,
or declarative programming in general,
with a single computation model and programming language
\cite{Hanus97DPLE}.

@@

In this paper, we present our practical results
with different search strategies implemented
in a new Curry system called
KiCS2 \cite{BrasselHanusPeemoellerReck11}.
KiCS2 compiles Curry programs into Haskell programs
where non-deterministic values or computations are represented
as tree structures so that flexible search strategies
can be supported.
Although the incomplete depth-first search strategy
is the most efficient one (provided that it is able
to find a result value),
we show that complete strategies,
like breadth-first search or iterative deepening search,
are a reasonable alternative that does not force
the programmer to consider the applied search strategy
in his program.

@@

A Curry program consists of the definition of data types and
operations on these types. 
Since the syntax of Curry is close to Haskell,
variables and function names usually
start with lowercase letters and the names of type and data constructors
start with an uppercase letter. The application of $f$
to $e$ is denoted by juxtaposition (``$f~e$'').
In addition, Curry allows free (logic) 
variables in conditions and right-hand sides of defining rules.
Note that in a functional logic language
operations might yield more than one result on the same input due to 
the logic programming features.
For instance, Curry contains a \emph{choice} operation defined by:
%
\begin{curry}
x ? _ = x
_ ? y = y
\end{curry}
%
The choice operation can be used to define other non-deterministic operations
like
%
\begin{curry}
coin = 0 ? 1  
\end{curry}
Thus, the expression \ccode{coin} has two values: \code{0} and \code{1}.
If expressions have more than one value, one wants to select
intended values according to some constraints,
typically in conditions of program rules.
A \emph{rule} has the form
\begin{curry}
$f~t_1\ldots{}t_n$ | $c$ = $e$
\end{curry}
where the (optional) condition $c$ is a \emph{constraint},
i.e., an expression of the built-in type
\code{Success}. For instance, the trivial constraint
\code{success} is a value of type \code{Success} that
denotes the always satisfiable constraint.
Thus, we say that a constraint $c$ is \emph{satisfied} if it
can be evaluated to \code{success}.
An \emph{equational constraint} $e_1 \,\code{=:=}\, e_2$ is satisfiable
if both sides $e_1$ and $e_2$ are reducible to unifiable values.
%Furthermore, if $c_1$ and $c_2$ are constraints,
%\code{$c_1\,$\&$\,\,c_2$} denotes their concurrent conjunction
%(i.e., both argument constraints are concurrently evaluated).

@@

An early approach
\cite{HanusSteiner98PLILP,Lux99FLOPS}
to encapsulating search in Curry
is based on a primitive search operator
\begin{curry}
try :: (a->Success) -> [a->Success]
\end{curry}
that takes a constraint abstraction, e.g., \code{(\bs{}x->x\,=:=\,coin)},
as input, evaluates it until the first non-deterministic step occurs,
and returns the result: the empty list in case of failure, a list with a single
element in case of success, or a list with at least two elements representing
a nondeterministic choice. For instance,
\code{try\,(\bs{}x->x\,=:=\,coin)} evaluates to
\code{[\bs{}x->x\,=:=\,0,\,\bs{}x->x\,=:=\,1]}.
Based on this primitive, one can define various search strategies
to explore the search space and return its solutions.
\cite{Lux99FLOPS} shows an implementation of this primitive.

Although typical search operators of Prolog,
like \code{findall}, \code{once}, or negation-as-failure,
can be implemented using \code{try},
it became also clear that the combination of
encapsulated search and demand-driven evaluation
and sharing causes further complications \cite{BrasselHanusHuch04JFLP}.
For instance, in an expression like
\begin{curry}
let y=coin in try (\x->x=:=y)
\end{curry}
it is not obvious whether the non-determinism caused by the evaluation
of \code{coin} (introduced outside
but demanded inside the search operator) should be encapsulated or not.
Hence, the result of this expression might depend on the evaluation order.
For instance,
if \code{coin} is evaluated before the \code{try} expression,
it results in two computations where \code{y} is bound to \code{0}
in one computation and to \code{1} in the other computation.
Hence, \code{try} does not encapsulate the nondeterminism of \code{coin}
(this is the semantics of \code{try} implemented in \cite{Lux99FLOPS}).
However, if \code{coin} is evaluated inside the capsule of \code{try}
(because it is not demanded before), then the nondeterminism of
\code{coin} is encapsulated.
These and more peculiarities are discussed in \cite{BrasselHanusHuch04JFLP}.
Furthermore, the order of the solutions might depend on the
textual order of program rules or the evaluation time
(e.g., in parallel implementations).
Therefore, \cite{BrasselHanusHuch04JFLP} contains a proposal
for another primitive search operator:
\begin{curry}
getSearchTree :: a -> IO (SearchTree a)
\end{curry}
It takes an expression and delivers a search tree representing
the search space when evaluating the input:
\begin{curry}
data SearchTree a = Value a
                  | Fail
                  | Or (SearchTree a) (SearchTree a)
\end{curry}
\code{(Value v)} and \code{Fail} represents a single value
or a failure (i.e., no value), respectively,
and \code{(Or t1 t2)} represents a choice (i.e., a non-deterministic value)
between two search trees \code{t1} and \code{t2}.
Since \code{getSearchTree} is an I/O action, its result (in particular,
the order of solutions) depends on the current environment, e.g.,
time of evaluation.
To avoid the complications w.r.t.\ shared variables,
\code{getSearchTree} implements a \emph{strong encapsulation view},
i.e., conceptually, the argument of \code{getSearchTree}
is cloned before the evaluation starts in order to cut any sharing
with the environment.
Furthermore, the structure of the search tree is computed lazily
so that an expression with infinitely many values does not cause
the nontermination of the search operator if one is interested in only
one solution.

@@

values []           = []
values (Fail:ts)    = values ts
values (Value x:ts) = x : values ts
values (Or _ _:ts)  = values ts

children []           = []
children (Fail:ts)    = children ts
children (Value _:ts) = children ts
children (Or x y:ts)  = x:y:children ts
\end{curry}
%
The operation \code{values} extracts the values
in each level of the tree and the operation \code{children}
extracts all direct ancestors of a level in order to recursively
collect their values.

@@

Another complete search strategy with a superior memory behavior
is iterative-deepening search.
Basically, it is depth-first search strategy with a depth-bound
which is incremented in each iteration.
Thus, we compute in each iteration a list of values
together with some information whether we have aborted
(due to the depth-bound) the computation of further possible values.
For this purpose, we define list structures that can also end
with an \code{Abort}:
%
\begin{curry}
data AbortList a = Nil | Cons a (AbortList a) | Abort
\end{curry}
%
and define the concatenation on such lists:
%
\begin{curry}
concA :: AbortList a -> AbortList a -> AbortList a
concA Abort       Abort       = Abort
concA Abort       Nil         = Abort
concA Abort       (Cons x xs) = Cons x (concA Abort xs)
concA Nil         ys          = ys
concA (Cons x xs) ys          = Cons x (concA xs ys)
\end{curry}
%
Now we define an operation to collect values in a search tree
within some level bounds (to avoid the repeated collection
of values found in each iteration):
%
\begin{curry}
collectInBounds :: Int -> Int -> SearchTree a -> AbortList a
collectInBounds oldbound newbound st = collectLevel newbound st
 where
  collectLevel _ Fail      = Nil
  collectLevel d (Value x) = if d <= newbound - oldbound
                             then Cons x Nil
                             else Nil
  collectLevel d (Or x y)  =
    if d>0
    then concA (collectLevel (d-1) x) (collectLevel (d-1) y)
    else Abort
\end{curry}
%
Now, the entire search strategy consists of
repeated calls to \code{collectInBounds} as long as the result list
is aborted. In order to experiment with different parameters,
the initial depth bound and the method to increase the depth bound
in each iteration are passed as parameters to the main operation:
%
\begin{curry}
allValuesIDS :: Int -> (Int -> Int) -> SearchTree a -> [a]
allValuesIDS initdepth incrdepth st =
  iterIDS initdepth (collectInBounds 0 initdepth st)
 where
  iterIDS _ Nil = []
  iterIDS n (Cons x xs) = x : iterIDS n xs
  iterIDS n Abort =
    let newdepth = incrdepth n
     in iterIDS newdepth (collectInBounds n newdepth st)
\end{curry}
%
The key advantage of depth-first search
in comparison to breadth-first search is its memory behavior:
whereas breadth-first search has to store all nodes
in a level of the search tree in parallel, depth-first search
only needs to store the nodes in the branch from the root
to the current node under investigation.
Since iterative deepening uses depth-first search in each iteration,
it should have a memory behavior similarly to depth-first search
and, in case of wide search trees,
better than breadth-first search.
The price for this behavior is the recomputation of the initial
goal in each iteration. 
However, since we defined iterative deepening on a search tree,
recomputation is not required. Instead, the
already evaluated part of the search tree is kept in memory.
Therefore, we also implemented the iterative deepening strategy
for top-level goals without an explicit search tree
but with a recomputation of the initial goal in each iteration
(see Section~\ref{sec:benchmarks}).

@@

In this section we evaluate the practical behavior
of the various search strategies discussed so far.
Since they are only supported by the Curry implementation KiCS2,
we use only this system for our evaluation.
A general comparison of KiCS2 and other Curry implementations
can be found in \cite{BrasselHanusPeemoellerReck11}.
Since KiCS2 compiles Curry programs into Haskell programs,
we used the Glasgow Haskell Compiler
(GHC 7.0.4, option -O2) to compile and execute
the generated target programs.
All benchmarks were executed on a 32bit Linux machine
running Ubuntu 11.10 with an Intel Core 2 Duo (2.13~GHz) processor and 4~GiB RAM.
The timings were performed with the time command measuring the
execution time (in seconds) of a compiled executable for each benchmark.
Due to the lack of precise measurements of the space behavior,
we measured only the execution times.

@@

Figure~\ref{fig:bench-all-solutions} shows the benchmark results
when all values of a given expression are computed
with various search strategies.
All benchmark programs are non-deterministic programs.
``PermSort'' sorts a list containing 15 elements by enumerating
all permutations and selecting the sorted ones,
``Last'' computes the last element \code{x}
of a list \code{xs} containing 10,000 elements
by solving the equation \ccode{ys++[x] =:= xs}
(see Section~\ref{sec:curry}),
``Half'' computes the half \code{y} of a natural number \code{x}
(in Peano representation) by solving the equation \code{y+y=:=x},
%``Embed'' searches for an embedding of a message
%which is ``hidden'' in another message
%(see \cite[Sect.~4.1]{Antoy10JSC}),
``Graph'' computes some path in a graph where the edges
are represented by a non-deterministic ``successor'' operation,
``HorseMan'' computes the numbers of horses and men
from given numbers of heads and feet by searching for
appropriate numbers in Peano representation,
``MAC'' solves the ``Missionaries and Cannibals'' puzzle
(see \cite[Sect.~3.1]{AntoyHanus02FLOPS})
for reasonable numbers of missionaries and cannibals
to obtain measurable run times,
and ``Queens'' computes safe placements of queens on a chess board.

@@

We have practically evaluated and compared
the efficiency of these strategies.
The benchmarks indicated that complete strategies
are a viable alternative to incomplete strategies,
that have been favored in the past due to limited memory requirements.
Thus, it could be reasonable to make
complete strategies the default in functional logic languages.
This would have a good impact on teaching declarative programming
to beginners.
Moreover, decoupling non-deterministic programs
from their search strategy could lead to a more declarative
programming style (instead of the use of predicates with
side effects as in Prolog) and enable more potential for
optimization, e.g., parallel search strategies.
Furthermore, if efficiency and memory limitations are important,
one can still use an efficient strategy like depth-first search
provided that it is able to find all solutions
(e.g., in case of a finite search space).
It is an interesting topic for future work
to statically approximate situations where
the use of incomplete strategies is sufficient.

>>>>

@@

\author{
Michael Hanus
\quad
Bj{\"o}rn Peem{\"o}ller
\quad
Fabian Reck \\
\\
Institut f\"ur Informatik, CAU Kiel, D-24098 Kiel, Germany \\
\texttt{\{mh|bjp|fre\}@informatik.uni-kiel.de}
}
\title{Search Strategies for Functional Logic Programming}
\begin{document}
\maketitle

@@

Functional logic languages combine the most important
features of functional and logic programming in a single language
(see \cite{AntoyHanus10CACM,Hanus07ICLP} for recent surveys).
In particular, they provide higher-order functions and demand-driven
evaluation from functional programming together with logic programming features
like non-deterministic search and computing with partial information
(logic variables).
This combination
led to new design patterns \cite{AntoyHanus02FLOPS,AntoyHanus11WFLP}
and better abstractions for application programming,
e.g., as shown for programming with databases
\cite{BrasselHanusMueller08PADL,Fischer05},
GUI programming \cite{Hanus00PADL},
web programming \cite{Hanus01PADL,Hanus06PPDP,HanusKoschnicke10PADL},
or string parsing \cite{CaballeroLopez99}.
Moreover, it is a good basis
to teach the ideas of functional and logic programming,
or declarative programming in general,
with a single computation model and programming language
\cite{Hanus97DPLE}.
The operational principles of functional logic languages
have also been used for other computation tasks,
like inverse computations \cite{AbramovGlueckKlimov06},
partial evaluation \cite{AlpuenteFalaschiVidal98TOPLAS}, or
generation of test cases \cite{FischerKuchen07,RuncimanNaylorLindblad08}.

@@

In this paper, we present our practical results
with different search strategies implemented
in a new Curry system called
KiCS2 \cite{BrasselHanusPeemoellerReck11}.
KiCS2 compiles Curry programs into Haskell programs
where non-deterministic values and computations are represented
as tree structures so that flexible search strategies
can be supported.
Although the incomplete depth-first search strategy
is the most efficient one (provided that it is able
to find a result value),
we show that complete strategies,
like breadth-first search or iterative deepening search,
are a reasonable alternative that does not force
the programmer to consider the applied search strategy
in his program.

@@

A Curry program consists of the definition of data types and
operations on these types. 
Since the syntax of Curry is close to Haskell,
variables and function names usually
start with lowercase letters and the names of type and data constructors
start with an uppercase letter. The application of $f$
to $e$ is denoted by juxtaposition (``$f~e$'').
In addition, Curry allows free (logic) 
variables in conditions and right-hand sides of defining rules.
Note that in a functional logic language
operations might yield more than one result on the same input due to 
the logic programming features.
For instance, Curry contains a \emph{choice} operation defined by:
%
\begin{curry}
x ? _ = x
_ ? y = y
\end{curry}
%
The choice operation can be used to define other non-deterministic operations
like
%
\begin{curry}
coin = 0 ? 1  
\end{curry}
Thus, the expression \ccode{coin} has two values: \code{0} and \code{1}.
If expressions have more than one value, one wants to select
intended values according to some constraints,
typically in conditions of program rules.
A \emph{rule} has the form
\begin{curry}
$f~t_1\ldots{}t_n$ | $c$ = $e$
\end{curry}
where the (optional) condition $c$ is a \emph{constraint},
i.e., an expression of the built-in type
\code{Success}. For instance, the trivial constraint
\code{success} is a value of type \code{Success} that
denotes the always satisfiable constraint.
Thus, we say that a constraint $c$ is \emph{satisfied} if it
can be evaluated to \code{success}.
An \emph{equational constraint} $e_1 \,\code{=:=}\, e_2$ is satisfiable
if both sides $e_1$ and $e_2$ are reducible to unifiable values.

@@

An early approach
\cite{HanusSteiner98PLILP,Lux99FLOPS}
to encapsulating search in Curry
is based on a primitive search operator
\begin{curry}
try :: (a->Success) -> [a->Success]
\end{curry}
that takes a constraint abstraction, e.g., \code{(\bs{}x->x\,=:=\,coin)},
as input, evaluates it until the first non-deterministic step occurs,
and returns the result: the empty list in case of failure, a list with a single
element in case of success, or a list with at least two elements representing
a non-deterministic choice. For instance,
\code{try\,(\bs{}x->x\,=:=\,coin)} evaluates to
\code{[\bs{}x->x\,=:=\,0,\,\bs{}x->x\,=:=\,1]}.
Based on this primitive, one can define various search strategies
to explore the search space and return its solutions.
\cite{Lux99FLOPS} shows an implementation of this primitive.

Although typical search operators of Prolog,
like \code{findall}, \code{once}, or negation-as-failure,
can be implemented using \code{try},
it became also clear that the combination of
encapsulated search and demand-driven evaluation
and sharing causes further complications \cite{BrasselHanusHuch04JFLP}.
For instance, in an expression like
\begin{curry}
let y=coin in try (\x->x=:=y)
\end{curry}
it is not obvious whether the non-determinism caused by the evaluation
of \code{coin} (introduced outside
but demanded inside the search operator) should be encapsulated or not.
Hence, the result of this expression might depend on the evaluation order.
For instance,
if \code{coin} is evaluated before the \code{try} expression,
it results in two computations where \code{y} is bound to \code{0}
in one computation and to \code{1} in the other computation.
Hence, \code{try} does not encapsulate the non-determinism of \code{coin}
(this is the semantics of \code{try} implemented in \cite{Lux99FLOPS}).
However, if \code{coin} is evaluated inside the capsule of \code{try}
(because it is not demanded before), then the non-determinism of
\code{coin} is encapsulated.
These and more peculiarities are discussed in \cite{BrasselHanusHuch04JFLP}.
Furthermore, the order of the solutions might depend on the
textual order of program rules or the evaluation time
(e.g., in parallel implementations).
Hence, it is difficult to define a search operator as a pure function.

Due to these considerations,
\cite{BrasselHanusHuch04JFLP} contains a proposal
for another primitive search operator:
\begin{curry}
getSearchTree :: a -> IO (SearchTree a)
\end{curry}
It takes an expression and delivers a search tree representing
the search space when evaluating the input:
\begin{curry}
data SearchTree a = Value a
                  | Fail
                  | Or (SearchTree a) (SearchTree a)
\end{curry}
\code{(Value v)} and \code{Fail} represent a single value
or a failure (i.e., no value), respectively,
and \code{(Or t1 t2)} represents a choice (i.e., a non-deterministic value)
between two search trees \code{t1} and \code{t2}.
Since \code{getSearchTree} is an I/O action, its result (in particular,
the order of subtrees) depends on the current environment, e.g.,
time of evaluation.
To avoid the complications w.r.t.\ shared variables,
\code{getSearchTree} implements a \emph{strong encapsulation view},
i.e., conceptually, the argument of \code{getSearchTree}
is cloned before the evaluation starts in order to cut any sharing
with the environment.
Furthermore, the structure of the search tree is computed lazily
so that an expression with infinitely many values does not cause
the nontermination of the search operator if one is interested in only
one solution.

@@

values []             = []
values (Fail    : ts) = values ts
values (Value x : ts) = x : values ts
values (Or _ _  : ts) = values ts

children []             = []
children (Fail    : ts) = children ts
children (Value _ : ts) = children ts
children (Or x y  : ts) = x : y : children ts
\end{curry}
%
The operation \code{values} extracts the values
in each level of the tree and the operation \code{children}
extracts all direct ancestors of a level in order to recursively
collect their values.

@@

Another complete search strategy with a superior memory behavior
is iterative-deepening search.
Basically, it is a depth-first search strategy with a depth-bound
which is incremented in each iteration.
Thus, we compute in each iteration a list of values
together with some information whether we have aborted
(due to the depth-bound) the computation of further possible values.
For this purpose, we define list structures that can also end
with an \code{Abort}:
%
\begin{curry}
data AbortList a = Nil | Cons a (AbortList a) | Abort
\end{curry}
%
and define the concatenation on such lists:
%
\begin{curry}
concA :: AbortList a -> AbortList a -> AbortList a
concA Abort       Abort       = Abort
concA Abort       Nil         = Abort
concA Abort       (Cons x xs) = Cons x (concA Abort xs)
concA Nil         ys          = ys
concA (Cons x xs) ys          = Cons x (concA xs ys)
\end{curry}
%
Now we define an operation to collect values in a search tree
within some level bounds (to avoid the repeated collection
of values found in each iteration):
%
\begin{curry}
collectInBounds :: Int -> Int -> SearchTree a -> AbortList a
collectInBounds oldbound newbound st = collectLevel newbound st
 where
  collectLevel _ Fail      = Nil
  collectLevel d (Value x) = if d <= newbound - oldbound
                             then Cons x Nil
                             else Nil
  collectLevel d (Or x y)  =
    if d>0
    then concA (collectLevel (d-1) x) (collectLevel (d-1) y)
    else Abort
\end{curry}
%
Now, the entire search strategy consists of
repeated calls to \code{collectInBounds} as long as the result list
is aborted. In order to experiment with different parameters,
the initial depth bound and the method to increase the depth bound
in each iteration are passed as parameters to the main operation:
%
\begin{curry}
allValuesIDS :: Int -> (Int -> Int) -> SearchTree a -> [a]
allValuesIDS initdepth incrdepth st =
  iterIDS initdepth (collectInBounds 0 initdepth st)
 where
  iterIDS _ Nil = []
  iterIDS n (Cons x xs) = x : iterIDS n xs
  iterIDS n Abort =
    let newdepth = incrdepth n
     in iterIDS newdepth (collectInBounds n newdepth st)
\end{curry}
%
The key advantage of depth-first search
in comparison to breadth-first search is its memory behavior:
whereas breadth-first search has to store all nodes
in a level of the search tree in parallel, depth-first search
only needs to store the nodes in the branch from the root
to the current node under investigation.
Since iterative deepening uses depth-first search in each iteration,
it should have a memory behavior similarly to depth-first search
and, in case of wide search trees,
better than breadth-first search.
The price for this behavior is the recomputation of the initial
goal in each iteration. 
However, since we defined iterative deepening on a search tree,
recomputation is not required. Instead, the
already evaluated part of the search tree is kept in memory.
Therefore, we also implemented the iterative deepening strategy
for top-level goals without an explicit search tree
but with a recomputation of the initial goal in each iteration
(see Section~\ref{sec:benchmarks}).

@@

In this section we evaluate the practical behavior
of the various search strategies discussed so far.
Since they are only supported by the Curry implementation KiCS2,
we use this system for our evaluation.
A general comparison of KiCS2 and other Curry implementations
can be found in \cite{BrasselHanusPeemoellerReck11}.
Since KiCS2 compiles Curry programs into Haskell programs,
we used the Glasgow Haskell Compiler
(GHC 7.0.4, option -O2) to compile and execute
the generated target programs.
All benchmarks were executed on a 32bit Linux machine
running Ubuntu 11.10 with an Intel Core 2 Duo (2.13~GHz) processor and 4~GiB RAM.
The timings were performed with the time command measuring the
execution time (in seconds) of a compiled executable for each benchmark.
Due to the lack of precise measurements of the space behavior,
we measured only the execution times.

@@

Figure~\ref{fig:bench-all-solutions} shows the benchmark results
when all values of a given expression are computed
with various search strategies.
All benchmark programs are non-deterministic programs.
``PermSort'' sorts a list containing 15 elements by enumerating
all permutations and selecting the sorted ones,
``Last'' computes the last element \code{x}
of a list \code{xs} containing 10,000 elements
by solving the equation \ccode{ys++[x] =:= xs}
(see Section~\ref{sec:curry}),
``Half'' computes the half \code{y} of a natural number \code{x}
(in Peano representation) by solving the equation \code{y+y=:=x},
``Graph'' computes some path in a graph where the edges
are represented by a non-deterministic ``successor'' operation,
``HorseMan'' computes the numbers of horses and men
from given numbers of heads and feet by searching for
appropriate numbers in Peano representation,
``MAC'' solves the ``Missionaries and Cannibals'' puzzle
(see \cite[Sect.~3.1]{AntoyHanus02FLOPS})
for reasonable numbers of missionaries and cannibals
to obtain measurable run times,
and ``Queens'' computes safe placements of queens on a chess board.

@@

We have practically evaluated and compared
the efficiency of these strategies.
The benchmarks indicated that complete strategies
are a viable alternative to incomplete strategies,
that have been favored in the past due to limited memory requirements.
Decoupling non-deterministic programs
from their search strategy could lead to a more declarative
programming style (instead of the use of predicates with
side effects as in Prolog) and enable more potential for
optimization, e.g., parallel search strategies.
The explicit definition of search strategies
has also been advocated for combining logic programs
with different constraint solvers
\cite{FrankHofstedtPepperReckmann06,Schulte97CP}.

Our results indicate that it could be reasonable to make
complete strategies the default in functional logic languages.
This would have a good impact on teaching declarative programming
to beginners.
Furthermore, if efficiency and memory limitations are important,
one can still use an efficient strategy, like depth-first search,
provided that it is able to find all solutions
(e.g., in case of a finite search space).
It is an interesting topic for future work
to statically approximate situations where
the use of theoretically incomplete strategies is sufficient.
